comgen-outline

### 2017-07-06 paper/comgen-paper.Rmd
light edits to discussion

All models were implemented in the probabilisitic programming language WebPPL [@dippl].
The following simulations were implemented in the Rmarkdown document that generated this paper text and figures, using the R package RWebPPL.
That document, as well as the other models, analysis, data and experiments used in this paper can be found at \url{https://github.com/mhtess/comgen-paper}.

@@

A column then appeared to the right of the animal names with a property in the header (e.g., "lays eggs").
Participants were asked to fill in each row with the percentage of members of each of the species that had the property by giving a number (e.g., "50\%").
Eight property--columns appeared in the table, and this whole procedure was repeated 2 times.
In total, each participant generated 10 animal names and reported on the prevalence of sixteen properties for 22 animals (their own 10 and the experimentally-supplied 12). 
<!-- For a full list of the properties, and generic sentences used in Expt. 1b, see Table 2 (Appendix). -->
The experiment can be viewed at \url{http://stanford.edu/~mtessler/experiments/generics/experiments/real-kinds/prior-2.html}.

@@

For the experimental trials, we anticipated there might be different beliefs about the frequency of events depending on whether the actor is male or female, so we asked about both genders. Participants answered both questions for both genders on each slide (4 questions total per slide, order of male / female randomized between-subjects), and every participant completed all 31 items in a randomized order.
The experiment can be viewed at \url{http://stanford.edu/~mtessler/habituals/experiments/priors/priors-2.html}.

@@

Participants were asked whether they agreed or disagreed with the corresponding habitual sentence: "\textsc{person does x}" (e.g., \emph{Bill smokes cigarettes}).
Participants completed 37 trials, which were composed of the 31 items from the prior elicitation task randomly paired with either a male or female character name.
6 of these items were then also paired with a name of the opposite gender (e.g., participants saw both a female character and a male character who watched professional football).
These were used for an exploratory analysis on differences in endorsements by gender of the target character.
The experiment can be viewed at \url{http://stanford.edu/~mtessler/habituals/experiments/truth-judgments/tj-2.html}.

@@

# fig.hab.freq.predictive.vs.past <- ggplot(d.hab.predictive.summary, 
#        aes(x=logAnnualPastRate, 
#                               y = mean, 
#                               ymin = ci_lower, ymax = ci_upper, 
#                               fill = condition, shape=condition))+
#   geom_jitter(position = position_jitter(width = .13), alpha = 0.6)+
#   scale_shape_manual(values=c(21,22,23))+
#   geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
#   # #geom_errorbar(width=0.1)+
#   scale_x_continuous(limits = c(-3,6), 
#                      breaks = c(-3, 0, 2.5, 4, 5.9),
#                      labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
#   scale_y_continuous(limits = c(-3,6), 
#                      breaks = c(-3, 0, 2.5, 4, 5.9),
#                      labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
#   guides(fill=F, shape=F)+
#   scale_fill_solarized()+
#   xlab("Past frequency (log scale)")+
#   ylab("Predicted frequency (log scale)")+
#   theme(axis.text.x = element_text(angle = 90))

@@

Participants were asked ``In the next \textsc{time window}, how many times do you think \textsc{person} does \textsc{event}?'', where the \textsc{time window} was the same as given in the \emph{past frequency statement}.
The experiment in full can be viewed at \url{http://stanford.edu/~mtessler/habituals/experiments/priors/predictive-1.html}.

@@

```{r habituals-predictive-model-run, cache = T}
items <- levels(factor(d.hab.predictive$item))
conditions <- levels(factor(d.hab.predictive$condition))
n_samples <- 500
rs.habituals.bda.predictive <- data.frame()

@@

```{r habituals-predictive-forwardSample}
rs.habituals.bda.predictive.summary <- rs.habituals.bda.predictive %>% 
  spread(Parameter, value) %>%
  rowwise() %>%
  mutate(
    logannualRate = rnorm(n = 1, mean = mu, sd = sigma)
  ) %>%
  gather(key, val, mu, sigma, logannualRate) %>%
  group_by(item, condition, key) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))
```

```{r habituals-predictive-bda-figure}
fig.hab.freq.predictive.vs.past <- left_join(
  rs.habituals.bda.predictive.summary %>% filter(key == "logannualRate"),
  unique(d.hab.predictive %>% 
    select(item, condition, logAnnualPastRate)
  )) %>% ggplot(., 
       aes(x=logAnnualPastRate, y = MAP, 
                              ymin = cred_lower, ymax = cred_upper, 
                              fill = condition, shape=condition))+
  geom_jitter(position = position_jitter(width = .13), alpha = 0.6)+
  scale_shape_manual(values=c(21,22,23))+
  geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
  # #geom_errorbar(width=0.1)+
  scale_x_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
  guides(fill=F, shape=F)+
  scale_fill_solarized()+
  xlab("Past frequency (log scale)")+
  ylab("Predicted frequency MAP (log scale)")+
  theme(axis.text.x = element_text(angle = 90))
```

### Modeling and results

We analyze this data by building a simple Bayesian model of the measurement so that it may be integrated seamlessly into a joint model with the endorsement data (as we have done for all previous case studies).
As we assumed for the prior elicitation task (Expt. 2a), we assume participant's responses for each event $i$ and condition $c$ are generated from a log-normal distribution with unknown mean and variance.
$$
\ln d_{ic} \sim \text{Gaussian}(\mu_{ic}, \sigma_{ic})
$$

Figure \ref{fig:habituals-predictive-figure} (left) shows the Maximum A-Posteriori predicted future frequency as a function of the past frequency given to the participant and the type of causal information given. 
We observe in the baseline condition that future frequency perfectly tracks past frequency (e.g., participants believe if a person smoked cigarettes 3 times last month, they will smoke cigarettes 3 times next month). 
This means that our model makes identical predictions for Expt. 2 whether the target is past frequency or expected future frequency (indicating, as expected, that we must look to the new data to distinguish these models).
Critically, we observe the preventative information appreciably decreasing and the enabling information slightly increasing predicted frequency (Figure \ref{fig:habituals-predictive-elicitation} left; blue and green dots).

@@

We confirm these observations using a linear mixed-effects model, predicting the log-transformed responses from the log-transormed past frequency and the experimental condition (baseline, preventative, enabling).
To account for participants and item variability in this analysis, we also include random effects of intercept and condition for both participants and items.
Confirming our predictions, the preventative information led to significantly lower predictions for future frequency, relative to the baseline condition (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","t value"],2)`$
).
There was also tendency for the enabling information to lead to higher predictions for future frequency, relative to baseline (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","t value"],2)`$
).
Finally, past frequency was a significant predictor of predicted future frequency (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","t value"],2)`$
).

@@

```{r habituals-predictive-rsa}
n_chains <- 3
# n_samples <- 500000
# burn <- n_samples / 2
# lg <- 150
model_prefix <- "results-habituals-predictive-jointModel-S1-smtncs_habitual-"
n_samples <- 100000
burn <- n_samples / 2
lg <- 20

m.hab.pred.samp <- data.frame()
m.hab.past.samp <- data.frame()

# for (i in seq(1, n_chains)){
#   mi <- fread(paste(project.path,  "models/habituals/results/",
#                     model_prefix, "targetH_predictiveMean-",
#                     n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   m.hab.pred.samp <- bind_rows(m.hab.pred.samp, mi %>% mutate(chain = i))

@@

#}
# 
# save(m.hab.pred.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "predictive-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
# 
# save(m.hab.past.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "past-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))

@@

                                 ))) %>%
  filter(src == "Speaker model - predictive frequency") %>%
  ggplot(., aes ( x = MAP, xmin = cred_lower, xmax = cred_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = condition))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 1)+
  #guides(fill = F)+
  #theme(legend.position = "bottom")+
  scale_fill_solarized()
```

```{r fig.height = 14, fig.width=10, echo = F, eval=F}
m.hab.pred.samp %>%
  #filter(type == "predictive", C == "predictiveFreq") %>%
  filter(type == "predictive", C == "endorsement") %>%
  mutate(E = as.numeric(as.character(E))) %>%
  ggplot(., aes(x = E, fill = D))+
  geom_histogram(position = position_dodge())+
  #facet_grid(B~E)+
  facet_wrap(~B)
```

```{r habituals-predictive-figure, fig.cap="A: Results of Expt 3a. B: Results of Expt. 3b as they relate to the predicted frequencies from Expt. 3a, C & D: Endorsement models based on past frequency (C) and predictive frequency (D).", fig.width = 9, fig.height = 9}
legend_b <- get_legend(habituals.predictive.predictive.model + theme(legend.position="bottom"))

@@

prow <- plot_grid(fig.hab.freq.predictive.vs.past, fig.hab.endorse.vs.predfreq, habituals.predictive.past.model, habituals.predictive.predictive.model + theme(legend.position="none"), labels = c("A", "B", "C", "D"), ncol = 2, align = 'v')

plot_grid( prow, legend_b, ncol = 1, rel_heights = c(5, 0))
```

@@

We recruited `r n.subj.hab.pred.endorse` participants from MTurk, using the same criterion as Expt. 2b.
The experiment took `r round(ave.seconds.hab.pred.endorse / 60)` minutes on average on participants were compensated \$0.40 for their work.
None of the participants had completed Expt. 2c.
The only difference from Expt. 2c is the dependent measure. 
On each trial, participants were asked if they agreed or disagreed with the corresponding habitual sentence (as in Expt. 2b).
The experiment can be viewed at \url{http://stanford.edu/~mtessler/habituals/experiments/truth-judgments/tj-3-preventative.html}.

@@

We use our formal model to test whether past or predictive frequency is what is being communicated by these statements.
To formalize the predictive frequency speaker model, we use the predictive frequency data (Expt. 2c) as the $h$ that the speaker model $S$ (Eq. \ref{eq:S1}) is aiming to communicate. 
The past frequency model is constructed using the past frequency supplied to participants as the $h$ that the speaker $S$ model is trying to communicate.

We analyze this model in the same Bayesian data analysis regime as for our previous models, integrating the simple Bayesian analysis of the predictive frequency data (Expt. 2c; for the predictive model only) with the joint model of the priors (Expt. 2a) and the new endorsement data (Expt. 2d).
We use the same priors over the parameters as before, and learn about the posterior distribution by collecting three independent MCMC chains of N iterations (removing the first M for burn-in). 
Figure \ref{fig:habituals-predictive-figure} C \& D shows the resulting model predictions for the past frequency and the predictive frequency speaker models. 
Participants judgments of the habitual statements was indeed influenced by the causal manipulations in the way predicted by the speaker model using the predictive frequency as input ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.predictive`$; MSE = $`r mse.habituals.rsa.predictive`$).
The model based on past frequency does not make different predictions for the different causal manipulation conditions and does a poor job at explaining the endorsements ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.past`$; MSE = $`r mse.habituals.rsa.past`$).

@@

Habitual language conveys generalizations about events and our model decides if a habitual sentence is a pragmatically useful way to describe the rate at which a person does an action, taking into account the listener’s prior beliefs about the action (measured in Expt. 2a). 
We have seen that our computational model endorses statements that communicate generalizations about events with the same sensitivity to context that people exhibit (Expt. 2b).
In Experiment 2b, we varied the type of event and the past frequency with which the person did the action, and found substantial graded variability in the endorsement judgments of people.
<!-- Naive baseline models could not explain these endorsements judgments and our communicative model provided the most parsimonious account of the data in comparison to non-Bayesian models.  -->
<!-- In Experiment 3, we looked at interpretations of habitual statements about different events and again found substantial gradedness that our theory provided the most parsimonious account of. -->
<!-- The theory defines a pair of interpretation and endorsement models and these models can simultaneously account for both data sets. -->
In Experiments 2c \& 2d, we further investigated the nature of the underlying “propensity” scale by introducing enabling and disabling causal evidence, measuring the predicted future frequency (Expt. 2c) and using that, with the model, to predict the felicity of habitual sentences (Expt. 2c). 
To our knowledge, the experiments presented here are first empirical investigations into the truth conditions of habitual sentences and the first test of a formal psychological model of generalizations in language.
In these experiments, we directly manipulated the frequency that the speaker was trying to communicate while measuring the propensity priors.
Thus, it is still possible that the propensity priors that are central to our communicative theory of generalizations are somehow epiphenomenal, correlational associated with generic endorsement but not directly causally related.
In our final case study, we extend our theory to communicating generlizations about causes, and test experimentally whether the propensity prior is indeed causally related to generic endorsement.

@@

Learning that an event consistently tends to result in another event serves as a foundation for an intuitive theory of causality [@Goodman2011; @Kemp2010a].
Communicating generalizations about causally related events is potentially highly useful because no one individual can run all the experiments that she wants to run in a life time. 
The history of civilizations has shown that it is incredibly useful to divide labor: Some members can investigate modifications to the food preparation procedures while others can examine different ways of cutting stone.

>>>>

All models were implemented in the probabilisitic programming language WebPPL [@dippl].
The following simulations were implemented in the RMarkdown document that generated this paper text and figures, using the R package RWebPPL.
That document, as well as the other models, analysis, data and links to experiments used in this paper can be found at \url{https://mhtess.github.io}.

@@

A column then appeared to the right of the animal names with a property in the header (e.g., "lays eggs").
Participants were asked to fill in each row with the percentage of members of each of the species that had the property by giving a number (e.g., "50\%").
Eight property--columns appeared in the table, and this whole procedure was repeated 2 times.
In total, each participant generated 10 animal names and reported on the prevalence of sixteen properties for 22 animals (their own 10 and the experimentally-supplied 12). 
<!-- For a full list of the properties, and generic sentences used in Expt. 1b, see Table 2 (Appendix). -->

@@

For the experimental trials, we anticipated there might be different beliefs about the frequency of events depending on whether the actor is male or female, so we asked about both genders. Participants answered both questions for both genders on each slide (4 questions total per slide, order of male / female randomized between-subjects), and every participant completed all 31 items in a randomized order.

@@

Participants were asked whether they agreed or disagreed with the corresponding habitual sentence: "\textsc{person does x}" (e.g., \emph{Bill smokes cigarettes}).
Participants completed 37 trials, which were composed of the 31 items from the prior elicitation task randomly paired with either a male or female character name.
6 of these items were then also paired with a name of the opposite gender (e.g., participants saw both a female character and a male character who watched professional football).
These were used for an exploratory analysis on differences in endorsements by gender of the target character.

@@

fig.hab.freq.predictive.vs.past <- ggplot(d.hab.predictive.summary,
       aes(x=logAnnualPastRate,
                              y = mean,
                              ymin = ci_lower, ymax = ci_upper,
                              fill = condition, shape=condition))+
  geom_jitter(position = position_jitter(width = .13), alpha = 0.6)+
  scale_shape_manual(values=c(21,22,23))+
  geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
  # #geom_errorbar(width=0.1)+
  scale_x_continuous(limits = c(-3,6),
                     breaks = c(-3, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-3,6),
                     breaks = c(-3, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
  guides(fill=F, shape=F)+
  scale_fill_solarized()+
  xlab("Past frequency (log scale)")+
  ylab("Predicted frequency (log scale)")+
  theme(axis.text.x = element_text(angle = 90))

@@

Participants were asked ``In the next \textsc{time window}, how many times do you think \textsc{person} does \textsc{event}?'', where the \textsc{time window} was the same as given in the \emph{past frequency statement}.

@@

```{r habituals-predictive-model-run, cache = T, eval=F}
items <- levels(factor(d.hab.predictive$item))
conditions <- levels(factor(d.hab.predictive$condition))
n_samples <- 5000
rs.habituals.bda.predictive <- data.frame()

@@

```{r eval = F}
  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, fill = condition))+
#  geom_density(size = 1, aes( y = ..scaled.. ))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
               position = position_dodge())+
  facet_wrap(~item, scales = 'free')+
  scale_color_solarized()+
  #scale_x_continuous(limits = c(-0.05,1.05), breaks = c(0, 0.5, 1)) +
  scale_x_continuous(limits = c(-5,10)) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r habituals-predictive-bda-forwardSample, fig.width=14, eval=F}
rs.habituals.bda.predictive.samples <- rs.habituals.bda.predictive %>% 
  spread(Parameter, value) %>%
  rowwise() %>%
  mutate(
    logannualRate = rnorm(n = 1, mean = mu, sd = sigma)
  )

  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, color = condition))+
#  geom_density(size = 1, aes( y = ..scaled.. ))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
               position = position_dodge())+
  facet_wrap(~item, scales = 'free')+
  scale_color_solarized()+
  #scale_x_continuous(limits = c(-0.05,1.05), breaks = c(0, 0.5, 1)) +
  scale_x_continuous(limits = c(-5,10)) +
  theme(strip.text.y = element_text(angle = 0))

bind_rows(
  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate),
  rs.habituals.bda.predictive.samples %>% 
    select(item, condition, logannualRate) %>%
    rename(logAnnualPredictiveRate = logannualRate) %>%
    mutate(src = 'model') 
  ) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, color = src))+
  geom_density(size = 1, aes( y = ..scaled.. ))+
#  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
#               position = position_dodge())+
  facet_grid(condition ~ item )+
  scale_color_solarized()+
  scale_x_continuous(limits = c(-5,10)) +
  #scale_y_continuous(limits = c(-0.01,1.01), breaks = c(0, 0.5, 1)) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r habituals-predictive-bda-summary, eval =F}
rs.habituals.bda.predictive.summary <- rs.habituals.bda.predictive %>% 
  spread(Parameter, value) %>%
  rowwise() %>%
  mutate(
    logannualRate = rnorm(n = 1, mean = mu, sd = sigma)
  ) %>%
  gather(key, val, mu, sigma, logannualRate) %>%
  group_by(item, condition, key) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))
```

```{r habituals-predictive-bda-figure, eval = F}
fig.hab.freq.predictive.vs.past <- left_join(
  rs.habituals.bda.predictive.summary %>% filter(key == "logannualRate"),
  unique(d.hab.predictive %>% 
    select(item, condition, logAnnualPastRate)
  )) %>% ggplot(., 
       aes(x=logAnnualPastRate, y = MAP, 
                              ymin = cred_lower, ymax = cred_upper, 
                              fill = condition, shape=condition))+
  geom_jitter(position = position_jitter(width = .13), alpha = 0.6)+
  scale_shape_manual(values=c(21,22,23))+
  geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
  # #geom_errorbar(width=0.1)+
  scale_x_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
  guides(fill=F, shape=F)+
  scale_fill_solarized()+
  xlab("Past frequency (log scale)")+
  ylab("Predicted frequency MAP (log scale)")+
  theme(axis.text.x = element_text(angle = 90))
```

<!-- We analyze this data by building a simple Bayesian model of the measurement so that it may be integrated seamlessly into a joint model with the endorsement data (as we have done for all previous case studies). -->
<!-- As we assumed for the prior elicitation task (Expt. 2a), we assume participant's responses for each event $i$ and condition $c$ are generated from a log-normal distribution with unknown mean and variance. -->
<!-- $$ -->
<!-- \ln d_{ic} \sim \text{Gaussian}(\mu_{ic}, \sigma_{ic}) -->
<!-- $$ -->

Figure \ref{fig:habituals-predictive-figure} (left) shows the mean predicted future frequency as a function of the past frequency given to the participant and the type of causal information given. 
We observe in the baseline condition that future frequency perfectly tracks past frequency (e.g., participants believe if a person smoked cigarettes 3 times last month, they will smoke cigarettes 3 times next month). 
This means that our model makes identical predictions for Expt. 2b whether the target is past frequency or expected future frequency (indicating, as expected, that we must look to the new data to distinguish these models).
Critically, we observe the preventative information appreciably decreasing and the enabling information slightly increasing predicted frequency (Figure \ref{fig:habituals-predictive-elicitation} left; blue and green dots).

@@

We confirm these observations using a linear mixed-effects model, predicting the log-transformed responses from the log-transormed past frequency and the experimental condition (baseline, preventative, enabling).
To account for participant and item variability in this analysis, we also include random effects of intercept and condition for both participants and items.
Confirming our predictions, the preventative information led to significantly lower predictions for future frequency, relative to the baseline condition (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionpreventative","t value"],2)`$
).
There was also tendency for the enabling information to lead to higher predictions for future frequency, relative to baseline (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["conditionenabling","t value"],2)`$
).
Finally, past frequency was a significant predictor of predicted future frequency (
$\beta = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","Estimate"],2)`$;
$SE = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","Std. Error"],2)`;$
$t = `r round(lmer.rs.hab.predictive.summary[["coefficients"]]["logAnnualPastRate","t value"],2)`$
).

@@

```{r habituals-predictive-rsa}
n_chains <- 3
# n_samples <- 500000
# burn <- n_samples / 2
# lg <- 150
model_prefix <- "results-habituals-predictive-jointModel-S1-smtncs_habitual-"
#model_prefix <- "results-habituals-predictive-jointModel-predMin-3-S1-smtncs_habitual-"
n_samples <- 100000
burn <- n_samples / 2
lg <- 20

m.hab.pred.samp <- data.frame()
m.hab.past.samp <- data.frame()
# m.hab.pred.samp <- fread(paste(project.path,
#                                 "models/habituals/results/",
#                                 "results-habituals-predictive-meanPredictive-min-5-jointModel-S1-smtncs_habitual-targetH_predictive-10000_burn5000_lag20_chain1.csv",
#                                 sep = ""))

# for (i in seq(1, n_chains)){
#   mi <- fread(paste(project.path,  "models/habituals/results/",
#                     model_prefix, "targetH_predictive-",
#                     n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   m.hab.pred.samp <- bind_rows(m.hab.pred.samp, mi %>% mutate(chain = i))

@@

# 
# save(m.hab.pred.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "predictive-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
# 
# save(m.hab.past.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "past-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))

@@

                                 ))) %>%
  filter(src == "Speaker model - predictive frequency") %>%
  ggplot(., aes ( x = MAP, xmin = cred_lower, xmax = cred_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = condition))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 1)+
  #guides(fill = F)+
  #theme(legend.position = "bottom")+
  scale_fill_solarized()
#habituals.predictive.predictive.model
```

```{r fig.height = 6, fig.width=14, echo = F, eval=F}
m.hab.pred.samp %>%
  filter(type == "predictive", C == "predictiveFreq", E == "sigma") %>%
  #filter(type == "predictive", C == "endorsement") %>%
#  mutate(E = as.numeric(as.character(E))) %>%
  ggplot(., aes(x = val, fill = D))+
  geom_histogram(position = position_dodge())+
  #facet_grid(B~E)+
  facet_grid(D~B)
```

```{r, eval = F}
m.hab.predfreq.summary <- m.hab.pred.samp %>%
  filter(type == "predictive", C == "predictiveFreq") %>%
  group_by(B, D, E) %>%
  summarize(MAP = estimate_mode(val)) %>% spread(E, MAP)

left_join(m.hab.predictive.endorse.rsa %>%
  mutate(src = factor(src, levels = c( 
                                      "past_model",
                                      "predictive_model"
                                     ),
                      labels = c(
                                 "Speaker model - past frequency",
                                 "Speaker model - predictive frequency"

                                 ))) %>%
  filter(src == "Speaker model - predictive frequency"), m.hab.predfreq.summary %>% rename(item = B, condition = D, freqmean = mean)) %>% View()
```

```{r habituals-predictive-figure, fig.cap="A: Results of Expt 3a. B: Results of Expt. 3b as they relate to the predicted frequencies from Expt. 3a, C & D: Endorsement models based on past frequency (C) and predictive frequency (D).", fig.width = 10.5, fig.height = 11}
legend_b <- get_legend(habituals.predictive.predictive.model + theme(legend.position="bottom"))

@@

grid.arrange(
  arrangeGrob(
    fig.hab.freq.predictive.vs.past, 
    fig.hab.endorse.vs.predfreq, 
    habituals.predictive.past.model, 
    habituals.predictive.predictive.model + theme(legend.position="none"),
    nrow=2, layout_matrix = cbind(c(1,3), c(2,4))),
legend_b, nrow=2,heights=c(8, 1))

#prow <- plot_grid(fig.hab.freq.predictive.vs.past, fig.hab.endorse.vs.predfreq, habituals.predictive.past.model, habituals.predictive.predictive.model #+ theme(legend.position="none"), labels = c("A", "B", "C", "D"), ncol = 2, align = 'v')

#plot_grid( prow, legend_b, ncol = 1, rel_heights = c(5, 0))
```

@@

We recruited `r n.subj.hab.pred.endorse` participants from MTurk, using the same criterion as Expt. 2b.
The experiment took `r round(ave.seconds.hab.pred.endorse / 60)` minutes on average on participants were compensated \$0.40 for their work.
None of the participants had completed Expt. 2c.
The only difference from Expt. 2c is the dependent measure. 
On each trial, participants were asked if they agreed or disagreed with the corresponding habitual sentence (as in Expt. 2b).

@@

We use our formal model to test whether past or predictive frequency is what is being communicated by these statements.
To formalize the predictive frequency speaker model, we use the mran predictive frequency data (Expt. 2c) as the $h$ that the speaker model $S$ (Eq. \ref{eq:S1}) is aiming to communicate. 
The past frequency model is constructed using the past frequency supplied to participants as the $h$ that the speaker $S$ model is trying to communicate.

We analyze this model in the same Bayesian data analysis regime as for our previous models.
We use the same priors over the parameters as before, and learn about the posterior distribution by collecting three independent MCMC chains of 100,000 iterations (removing the first 50,000 for burn-in). 
Figure \ref{fig:habituals-predictive-figure} (bottom, left and right) shows the resulting model predictions for the past frequency and the predictive frequency speaker models. 
Participants judgments of the habitual statements was indeed influenced by the causal manipulations in the way predicted by the speaker model using the predictive frequency as input ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.predictive`$; MSE = $`r mse.habituals.rsa.predictive`$).
The model based on past frequency does not make different predictions for the different causal manipulation conditions and does a poor job at explaining the endorsements ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.past`$; MSE = $`r mse.habituals.rsa.past`$).

@@

Habitual language conveys generalizations about events and our model decides if a habitual sentence is a pragmatically useful way to describe the rate at which a person does an action, taking into account the listener’s prior beliefs about the action (measured in Expt. 2a). 
We saw that our computational model endorses statements that communicate generalizations about events with the same sensitivity to context and frequency that people exhibit (Expt. 2b).
In Experiment 2b, we varied the type of event and the past frequency with which the person did the action, and found substantial graded variability in the endorsement judgments of people.
By manipulating (rather than measuring) the target frequency with which a person does an action, we showed how alternative models were insufficient to account for this gradience.
In particular, a fixed threshold model that has access to the same prior knowledge as our model and goes through the same information-theoretic decision making procedure as our speaker model is unable to make different predictions for different frequencies.
Only our uncertain threshold model was able to precisely account for the wide range of endorsements.

<!-- Naive baseline models could not explain these endorsements judgments and our communicative model provided the most parsimonious account of the data in comparison to non-Bayesian models.  -->
<!-- In Experiment 3, we looked at interpretations of habitual statements about different events and again found substantial gradedness that our theory provided the most parsimonious account of. -->
<!-- The theory defines a pair of interpretation and endorsement models and these models can simultaneously account for both data sets. -->
In Experiments 2c \& 2d, we further investigated the nature of the underlying “propensity” scale by introducing enabling and disabling causal evidence, measuring the predicted future frequency (Expt. 2c) and using that, with the model, to predict the felicity of habitual sentences (Expt. 2d). 
To our knowledge, the experiments presented here are first empirical investigations into the truth conditions of habitual sentences and the first test of a formal psychological model of such generalizations in language.

In these experiments, we directly manipulated the frequency that the speaker was trying to communicate while still measuring the priors over the propensity of the event.
Thus, it is still possible that the propensity priors that are central to our communicative theory of generalizations are somehow epiphenomenal, correlationally associated with generic endorsement but not directly causally related.
In our final case study, we extend our theory to communicating generlizations about causes, and experimentally test whether the propensity prior is indeed causally related to generic endorsement.

@@

Learning that an event consistently tends to result in another event serves as a foundation for an intuitive theory of causality [@Goodman2011; @Kemp2010a].
Communicating generalizations about causally related events is potentially highly valuable because relevant data can be prohibitively expensive and no one individual can run all the experiments that she wants to run in a life time. 
The history of civilizations has shown that it is incredibly useful to divide labor: Some members can investigate modifications to the food preparation procedures while others can examine different ways of cutting stone.

Our framework for communicating generalizations makes predictions about statements conveying genericity about a category of causal events. 
In this domain, an instance of the category is a single causal event, or *token causality* (e.g., adding a particular amount of a particular kind of yeast to an actual piece of dough). 
<!-- [Gerstenberg] -->
Ascribing causality to a singular pair of conjoined events is itself a challenging problem [e.g., @Gerstenberg2015how]. 
Inducing the existence of an abstract cause from observational evidence also depends in complex ways on an observer's background theory of causality [e.g., @Griffiths2005; @Schulz2008a].

Language provides a way to bypass the observational data needed to make a causal induction.
Indeed, causal language has been shown to faciliate causal learning for nonobvious causal relations in 2-year-olds [@Bonawitz2010].
